---
title: "WUP_04"
author: "Paulo"
date: "fall 2021"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
mydata<-read.csv(file = "GSScsv.csv")
```

```{r}
boxplot(WORDSUM~DEGREE,data=mydata)
```
The middle box and the far right box can go completely under the box to the left named "graduates". The means from the 2 sets are quite different from each other. The middle box has nothing in common with the middle 50% of the grad school box normally that means that there is a big different from the mean of the grad school students. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
mydata<-read.csv(file = "GSScsv.csv")
```

```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```
The people who junior college level education, a bachelors degree and and a grad school degree has a higher mean results than those compared to junior high school and high school category. As per the graph above, the mean for the degree holders area above 9.204545 while the non degree holders have a mean of less than 8.282353

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```
The data shows that the degree holders have a higher mean that those without. The p value on the other hand shows the opposite. In conclusion we have to reject the null hypothesis for this. The p value represents the chances of getting the results for this particular problem is very small. 


```{r}
par(mar = c(3, 16, 2, 2))
plot(TukeyHSD(model1,conf.level=.95),las=2)
```
We conclude that high school and graduate are different from each other and LH high school and graduates are also different.  As per graph above, LT highschool, graduate and highschool graduate do not hit the 0.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

